### 2022-03-03 Shp/Biodiv_features/IMMAs/iucn-imma.qmd
Data
Adding data folder except shipping

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.10.4-A Coruña">
  <identifier>iucn-imma.shp</identifier>
  <parentidentifier>
Important Marine Mammal Area (IMMA)</parentidentifier>
  <language>GBR</language>
  <type>dataset</type>
  <title>IUCN Joint SSC/WCPA Marine Mammal Protected Area
Task Force (IUCN MMPATF) Spatial Distribution of
Important Marine Mammal Areas (IMMA)</title>
  <abstract>The IUCN IMMA is a joint project between the IUCN Species Survival Commission (SSC) and World Commission on Protected Areas (WCPA). The IUCN IMMA layer indicates discrete portions of habitat, important to marine mammal species that have the potential to be delineated and managed for conservation. IMMAs consist of areas that may merit place-based protection and/or monitoring. IMMAs can be seen as a marine mammal layer indicative of biodiversity and potentially ecosystem health for consideration by governments, intergovernmental organisations, conservation groups, and the general public. 
The IUCN IMMA layer contains information, which describes indicates discrete portions of habitat, important to marine mammal species that have the potential to be delineated and managed for conservation. Information described include the name, location, marine mammal species of importance, the broad level and finer sub-divisions criteria used, and the area size of the IMMA.

Purpose of creation: The IUCN IMMA layer has been created to be compatible with and inform the identification of other area-based management tools and conservation prioritisation areas and include but are not restricted to Ecologically or Biologically Significant Area (EBSA), Key Biodiversity Area (KBA), Important Bird and Biodiversity Area (IBA), RAMSAR Convention Site (RAMSAR), Cetacean Critical Habitat (CCH), Biologically Important Area (BIA). IMMAs consist of areas that may merit place-based protection and/or monitoring and may be considered as a marine mammal layer indicative of biodiversity and potentially ecosystem health for consideration by governments, intergovernmental organisations, conservation groups, and the general public.

Creation methodology: This dataset is drawn from knowledge provided directly to the IUCN MMPATF IMMA Secretariat by experts but may also include some data consolidated from 3rd party sources where other information may not have been available. These areas have been assessed by regional experts and were further assessed by a panel of  independent reviewers. Where possible the IUCN MMPATF has  included a polygon for each IMMA sub-region, which has been independently assessed across separate regional expert identification workshops (dated 31.10.21). Guidance on the use of selection criteria for the identification of IMMAs is available for download via https://www.marinemammalhabitat.org/downloads/ </abstract>
  <keywords vocabulary="gmd:topicCategory">
    <keyword>Biota</keyword>
    <keyword>Environment</keyword>
    <keyword>Oceans</keyword>
  </keywords>
  <contact>
    <name>Dr. Michael J. Tetley</name>
    <organization>IUCN Joint SSC/WCPA Marine Mammal Protected Area Task Force
(IUCN MMPATF)</organization>
    <position>IMMA Data Coordinator</position>
    <voice>+44 1688 400 524</voice>
    <fax>NA</fax>
    <email>michael.tetley@imma-network.org</email>
    <role>distributor</role>
  </contact>
  <links>
    <link description="IMMA Website and eAtlas" url="https://www.marinemammalhabitat.org/imma-eatlas/" type="information" format="Website" mimeType="NA" name="IMMA eAtlas" size="11.1mb"/>
  </links>
  <history>Note that the map shown in the metadata sheet is using the October 2021 release. The IMMA e-Atlas shows the up to date and additional layers (e.g. areas of interest - AoI, candidate IMMAs - cIMMAs) not available for public download (https://www.marinemammalhabitat.org/imma-eatlas).

The IUCN IMMA dataset is not necessarily a complete representation of all the important areas that have been identified for a region; the quality of the IUCN IMMA depends on the accessibility of accurate, comprehensive, up-to-date expert information from data and knowledge holders.

Other regions not currently assessed within the IUCN IMMA will be addressed in future regional expert workshops. New information will be assessed on a decadal basis to determine new IMMAs in the region or re-examine existing IMMAs. Data are updated on a 6-12 monthly basis.
</history>
  <fees></fees>
  <constraints type="Use">IUCN MMPATF (2019) Global Dataset of Important Marine Mammal
Areas (IUCN IMMA). [Insert month/year of the version downloaded].
Made available under agreement on terms and conditions of use by
the IUCN Joint SSC/WCPA Marine Mammal Protected Areas Task
Force and accessible via the IMMA e-Atlas
http://www.marinemammalhabitat.org/imma-eatlas/</constraints>
  <constraints type="Use">IUCN MMPATF (2019) Global Dataset of Important Marine Mammal
Areas (IUCN IMMA). [Insert month/year of the version downloaded].
Made available under agreement on terms and conditions of use by
the IUCN Joint SSC/WCPA Marine Mammal Protected Areas Task
Force and accessible via the IMMA e-Atlas
http://www.marinemammalhabitat.org/imma-eatlas/</constraints>
  <rights>IUCN IMMA User Licence Agreement for the purposes of Scientific Research, Education, or Conservation. Please see the IUCN IMMA User Licence Agreement for full terms and conditions of use and
supply. Please contact the IMMA Data Coordinator for further conditions of commercial use michael.tetley@imma-network.org
</rights>
  <license>Attribution-NonCommercial
CC BY-NC</license>
  <encoding></encoding>
  <crs>
    <spatialrefsys>
      <wkt>GEOGCRS["WGS 84",DATUM["World Geodetic System 1984",ELLIPSOID["WGS 84",6378137,298.257223563,LENGTHUNIT["metre",1]]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],CS[ellipsoidal,2],AXIS["geodetic latitude (Lat)",north,ORDER[1],ANGLEUNIT["degree",0.0174532925199433]],AXIS["geodetic longitude (Lon)",east,ORDER[2],ANGLEUNIT["degree",0.0174532925199433]],USAGE[SCOPE["unknown"],AREA["World"],BBOX[-90,-180,90,180]],ID["EPSG",4326]]</wkt>
      <proj4>+proj=longlat +datum=WGS84 +no_defs</proj4>
      <srsid>3452</srsid>
      <srid>4326</srid>
      <authid>EPSG:4326</authid>
      <description>WGS 84</description>
      <projectionacronym>longlat</projectionacronym>
      <ellipsoidacronym>EPSG:7030</ellipsoidacronym>
      <geographicflag>true</geographicflag>
    </spatialrefsys>
  </crs>
  <extent>
    <spatial maxx="180" maxz="0" dimensions="2" minz="0" crs="EPSG:4326" maxy="45.79455600000000004" miny="-78.68547588799999914" minx="-180"/>
    <temporal>
      <period>
        <start>2016-09-30T20:00:00Z</start>
        <end>2021-10-31T11:06:24Z</end>
      </period>
    </temporal>
  </extent>
</qgis>